---
title: "Code Analysis"
output: pdf_document
date: "2024-05-14"
---

```{r setup, include=FALSE}
require(httr)
require(jsonlite)
require(dplyr)
library(readr)
library(janitor)
library(sf)
library(stringr)
library(readODS)
library(readxl)
library(janitor)
library(dplyr)
library(tidyr)
library(purrr)
library(tmap)
library(spdep)
library(spatialreg)
library(ggplot2)
library(plotly)
library(broom)
```


```{r read in the csv data }
london_data <- read_csv("Data/london_data.csv")
full_data <- read_csv("Data/full_data.csv") 
```


## 1) Proportion of land by land-use-type across England

```{r}
# Calculate total land area and proportions, handling NAs properly
total_land_area <- sum(full_data$total_developed_use, na.rm = TRUE) +
  sum(full_data$total_non_developed_use, na.rm = TRUE) +
  sum(full_data$total_vaccant_land_use, na.rm = TRUE)

proportions <- c(
  sum(full_data$total_developed_use, na.rm = TRUE),
  sum(full_data$total_non_developed_use, na.rm = TRUE),
  sum(full_data$total_vaccant_land_use, na.rm = TRUE)
) / total_land_area

# Assign names to the proportions for clarity
names(proportions) <- c("total_developed_use", "total_non_developed_use", "total_vacant_land_use")

# Print the results
print(proportions)
```

total_developed_use       total_non_developed_use   total_vacant_land_use 
            0.086874274             0.911214461             0.001911265 


## 2) Proportion of land by land-use-type across London

```{r}
# Calculate total land area and proportions, handling NAs properly
total_land_area <- sum(london_data$total_developed_use, na.rm = TRUE) +
  sum(london_data$total_non_developed_use, na.rm = TRUE) +
  sum(london_data$total_vaccant_land_use, na.rm = TRUE)

proportions <- c(
  sum(london_data$total_developed_use, na.rm = TRUE),
  sum(london_data$total_non_developed_use, na.rm = TRUE),
  sum(london_data$total_vaccant_land_use, na.rm = TRUE)
) / total_land_area

# Assign names to the proportions for clarity
names(proportions) <- c("total_developed_use", "total_non_developed_use", "total_vacant_land_use")

# Print the results
print(proportions)


# Summarizing land use types into a smaller dataframe for plotting
land_use_summary <- full_data %>%
  summarise(TotalDeveloped = sum(total_developed_use, na.rm = TRUE),
            TotalNonDeveloped = sum(total_non_developed_use, na.rm = TRUE),
            TotalVacant = sum(total_vaccant_land_use, na.rm = TRUE)) %>%
  pivot_longer(cols = everything(), names_to = "LandUseType", values_to = "Area")

# Pie chart using ggplot2
ggplot(land_use_summary, aes(x = "", y = Area, fill = LandUseType)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  theme_void() +
  labs(fill = "Land Use Type", title = "Proportion of Land by Land Use Types in England 2022", 
       caption = "Data collected from Land Use and Land Use Change Statistics 2022")

## what about for london
# Summarizing land use types into a smaller dataframe for plotting
land_use_summary2 <- london_data %>%
  summarise(TotalDeveloped = sum(total_developed_use, na.rm = TRUE),
            TotalNonDeveloped = sum(total_non_developed_use, na.rm = TRUE),
            TotalVacant = sum(total_vaccant_land_use, na.rm = TRUE)) %>%
  pivot_longer(cols = everything(), names_to = "LandUseType", values_to = "Area")

# Pie chart using ggplot2
ggplot(land_use_summary2, aes(x = "", y = Area, fill = LandUseType)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  theme_void() +
  labs(fill = "Land Use Type", title = "Proportion of Land by Land Use Types in London 2022", 
       caption = "Data collected from Land Use and Land Use Change Statistics 2022")


# Combine data for plotting
combined_data <- bind_rows(
  land_use_summary %>% mutate(Region = "England"),
  land_use_summary2 %>% mutate(Region = "London")
)


# Combine data for plotting
combined_data <- bind_rows(
  land_use_summary %>% mutate(Region = "England"),
  land_use_summary2 %>% mutate(Region = "London")
)

# Plotting both charts in a single image using a colorblind-friendly palette
ggplot(combined_data, aes(x = "", y = Area, fill = LandUseType)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  scale_fill_brewer(palette = "Set2", name = "Land Use Type") +
  facet_wrap(~Region, ncol = 1, scales = "free_y") +
  theme_void() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5),
    plot.caption = element_text(hjust = 0.5)
  ) +
  labs(title = "Proportion of Land by Land Use Types",
       caption = "Data: Land Use and Land Use Change Statistics 2022")
```

total_developed_use total_non_developed_use   total_vacant_land_use 
            0.406250021             0.586207775             0.007542204
            
            

## 1) Regional breakdown of land by land-use-type across England

```{r}
# Summarize total land use by type for each region
summary_table <- full_data %>%
  group_by(rgn21nm) %>%
  summarise(
    total_developed_use = sum(total_developed_use, na.rm = TRUE),
    total_non_developed_use = sum(total_non_developed_use, na.rm = TRUE),
    total_vaccant_land_use = sum(total_vaccant_land_use, na.rm = TRUE)
  )

# Print the summary table
print(summary_table)

# Optionally, create a bar plot for visual comparison
data_long <- pivot_longer(summary_table, cols = c("total_developed_use", "total_non_developed_use", "total_vaccant_land_use"), names_to = "Land_Use_Type", values_to = "Area")
ggplot(data_long, aes(x = rgn21nm, y = Area, fill = Land_Use_Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Regional Breakdown of Land Use 2022", x = "Region", y = "Area (hectares)") +
  theme_bw()

# Calculate proportions
summary_table <- full_data %>%
  group_by(rgn21nm) %>%
  summarise(
    total_developed_use = sum(total_developed_use, na.rm = TRUE),
    total_non_developed_use = sum(total_non_developed_use, na.rm = TRUE),
    total_vacant_land_use = sum(total_vaccant_land_use, na.rm = TRUE),
    total_land_use = sum(total_developed_use + total_non_developed_use + total_vaccant_land_use, na.rm = TRUE)
  ) %>%
  mutate(
    proportion_developed = total_developed_use / total_land_use,
    proportion_non_developed = total_non_developed_use / total_land_use,
    proportion_vacant = total_vacant_land_use / total_land_use
  )

# Ensure column names are consistent
summary_table_long <- summary_table %>%
  pivot_longer(
    cols = c(proportion_developed, proportion_non_developed, proportion_vacant),
    names_to = "land_use_type",
    values_to = "proportion",
    names_prefix = "proportion_"
  )

# Correct the fill argument if necessary
ggplot(summary_table_long, aes(x = rgn21nm, y = proportion, fill = land_use_type)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Proportion of Land Use Types by Regions in England 2022",
       x = "Region",
       y = "Proportion",
       fill = "Land Use Type", 
       caption = "Data Source: Land use and land use statistics 2022") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# ggplot(summary_table_long, aes(x = rgn21nm, y = proportion, fill = land_use_type)) +
#   geom_bar(stat = "identity", position = "dodge", color = "black", size = 0.5) +
#   labs(title = "Proportional Regional Breakdown of Land Use 2022", 
#        x = "Region", 
#        y = "Proportion",
#        fill = "Land Use Type") +
#   scale_y_continuous(labels = percent_format()) +
#   scale_fill_brewer(palette = "Set2") +  # Using a professional color palette
#   theme_classic() +
#   theme(
#     text = element_text(family = "Courier", size = 12),
#     plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
#     axis.title.x = element_text(size = 12),
#     axis.title.y = element_text(size = 12),
#     axis.text.x = element_text(angle = 45, hjust = 1),
#     legend.title = element_text(size = 12),
#     legend.text = element_text(size = 10),
#     panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "gray"),
#     panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "lightgray")
#   )

```



#Regional breakdown of land use categories within developed land and undeveloped land in England

```{r}
#develeoped
developed_data <- full_data %>%
  group_by(rgn21nm) %>%
  summarise(
    CommunityServices = sum(group_community_services, na.rm = TRUE),
    Defence = sum(group_defence, na.rm = TRUE),
    IndustryCommerce = sum(group_industry_commerce, na.rm = TRUE),
    MineralsLandfill = sum(group_minerals_landfill, na.rm = TRUE),
    OtherDeveloped = sum(group_other_developed, na.rm = TRUE),
    Residential = sum(group_residential, na.rm = TRUE),
    TransportUtilities = sum(group_transport_utilities, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = -rgn21nm, names_to = "LandUseType", values_to = "Area")

ggplot(developed_data, aes(x = rgn21nm, y = Area, fill = LandUseType)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Regional Breakdown of Developed Land Use Types in England 2022", x = "Region", y = "Area (hectares)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#non-developed
non_developed_data <- full_data %>%
  group_by(rgn21nm) %>%
  summarise(
    Agriculture = sum(group_agriculture, na.rm = TRUE),
    Forestry = sum(group_forestry, na.rm = TRUE),
    OutdoorRecreation = sum(group_outdoor_recreation, na.rm = TRUE),
    ResidentialGardens = sum(group_residential_gardens, na.rm = TRUE),
    UndevelopedLand = sum(group_undeveloped_land, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = -rgn21nm, names_to = "LandUseType", values_to = "Area")

ggplot(non_developed_data, aes(x = rgn21nm, y = Area, fill = LandUseType)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Regional Breakdown of Non-Developed Land Use Types in England 2022", x = "Region", y = "Area (hectares)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


## #Regional breakdown of land use categories within developed land and undeveloped land in London
 
```{r}
#developed
developed_data <- london_data %>%
  group_by(rgn21nm) %>%
  summarise(
    CommunityServices = sum(group_community_services, na.rm = TRUE),
    Defence = sum(group_defence, na.rm = TRUE),
    IndustryCommerce = sum(group_industry_commerce, na.rm = TRUE),
    MineralsLandfill = sum(group_minerals_landfill, na.rm = TRUE),
    OtherDeveloped = sum(group_other_developed, na.rm = TRUE),
    Residential = sum(group_residential, na.rm = TRUE),
    TransportUtilities = sum(group_transport_utilities, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = -rgn21nm, names_to = "LandUseType", values_to = "Area")

# Plot developed land use data
ggplot(developed_data, aes(x = rgn21nm, y = Area, fill = LandUseType)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Regional Breakdown of Developed Land Use Types in London 2022", x = "Region", y = "Area (hectares)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


## non- developed 
non_developed_data <- london_data %>%
  group_by(rgn21nm) %>%
  summarise(
    Agriculture = sum(group_agriculture, na.rm = TRUE),
    Forestry = sum(group_forestry, na.rm = TRUE),
    OutdoorRecreation = sum(group_outdoor_recreation, na.rm = TRUE),
    ResidentialGardens = sum(group_residential_gardens, na.rm = TRUE),
    UndevelopedLand = sum(group_undeveloped_land, na.rm = TRUE)
  ) %>%
  pivot_longer(cols = -rgn21nm, names_to = "LandUseType", values_to = "Area")

# Plot non-developed land use data
ggplot(non_developed_data, aes(x = rgn21nm, y = Area, fill = LandUseType)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Regional Breakdown of Non-Developed Land Use Types in London 2022", x = "Region", y = "Area (hectares)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
#### Lets map the infrascutre data

```{r}
# Example with 'community_buildings' and 'count'
tm_shape(london_data) +
  tm_polygons("cmmnty_",  
              style = "quantile", 
              palette = "Blues", 
              title = "Community Buildings (Hectares)") +
  tm_layout(title = "Distribution of Community Buildings by LSOA")


tm_shape(london_data) +
  tm_polygons("count", 
              style = "quantile", 
              palette = "Reds", 
              title = "Crime Counts") +
  tm_polygons("resdntl",
              style = "quantile", 
              palette = "Blues", 
              title = "Community Buildings (Hectares)") +
  tm_layout(title = "Crime Counts and Community Buildings by LSOA")

tm_shape(london_data) +
  tm_polygons("count", 
              style = "quantile", 
              palette = "Reds", 
              title = "Crime Counts") +
  tm_polygons("cmmnty_",
              style = "quantile", 
              palette = "Blues", 
              title = "Community Buildings (Hectares)") +
  tm_layout(title = "Crime Counts and Community Buildings by LSOA")
```



## correlation matrix

```{r}
install.packages("corrplot")
library(corrplot)

cor_matrix_dev <- cor(full_data[, c("group_community_services", "group_defence", "group_industry_commerce", 
                      "group_minerals_landfill", "group_other_developed", "group_residential", 
                      "group_transport_utilities")], use = "complete.obs")

# Plotting the correlation matrix
corrplot(cor_matrix_dev, method = "circle", type = "upper", order = "hclust",
         tl.col = "black", tl.srt = 45, addCoef.col = "black")

# Adding a title
title("Correlation Matrix for Developed Land Use Types")

cor_matrix_undev <- cor(full_data[, c("group_agriculture", "group_forestry", "group_outdoor_recreation", 
                        "group_residential_gardens", "group_undeveloped_land")], use = "complete.obs")

# Plotting the correlation matrix
corrplot(cor_matrix_undev, method = "circle", type = "upper", order = "hclust",
         tl.col = "black", tl.srt = 45, addCoef.col = "black")

# Adding a title
title("Correlation Matrix for Undeveloped Land Use Types")
```


### Lets explore the relationship with crime 

```{r}
##read in the sf objects here
full_data <- st_read("Data/full_data.shp")
london_data <- st_read("Data/london_data.shp")
london_data_by_crime <- st_read("Data/london_data_by_crime.shp")

# Scatter plot for crime counts and residential land use
ggplot(full_data, aes(x = ttl_dv_, y = count)) +
  geom_point(aes(color = rgn21nm), alpha = 0.5) +  # Optional: color points by region if relevant
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(title = "Relationship between developed Land Use and Crime Counts in England",
       x = "Residential Land Use (hectares)", y = "Crime Counts") +
  theme_minimal()

ggplot(full_data, aes(x = ttl_n__, y = count)) +
  geom_point(aes(color = rgn21nm), alpha = 0.5) +  # Optional: color points by region if relevant
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(title = "Relationship between undeveloped Land Use and Crime Counts in England",
       x = "Residential Land Use (hectares)", y = "Crime Counts") +
  theme_minimal()

# Reshape the data to long format for easier plotting
full_data_long <- full_data %>%
  pivot_longer(cols = c(ttl_dv_, ttl_n__), names_to = "LandUseType", values_to = "LandUse") %>%
  mutate(LandUseType = recode(LandUseType, ttl_dv_ = "Developed", ttl_n__ = "Undeveloped"))

# Plotting both regression lines on the same plot
ggplot(full_data_long, aes(x = LandUse, y = count, color = LandUseType)) +
  geom_point(alpha = 0.5) +  # You can still color points by region by adding: aes(group = rgn21nm)
  geom_smooth(method = "lm", se = TRUE, aes(linetype = LandUseType)) +  # Different linetypes for clarity
  labs(title = "Relationship between Land Use and Crime Counts in England",
       x = "Land Use (hectares)", y = "Crime Counts",
       color = "Land Use Type", linetype = "Land Use Type") +
  theme_minimal()

#Regression model 
model_dev_undev_eng <- lm(count ~ ttl_dv_ + ttl_n__, data = full_data)
summary(model_dev_undev_eng)

```


```{r}
## FOR LONDON 
ggplot(london_data, aes(x = ttl_dv_, y = count)) +
  geom_point(aes(color = rgn21nm), alpha = 0.5) +  # Optional: color points by region if relevant
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(title = "Relationship between developed Land Use and Crime Counts in London",
       x = "Residential Land Use (hectares)", y = "Crime Counts") +
  theme_minimal()

ggplot(london_data, aes(x = ttl_n__, y = count)) +
  geom_point(aes(color = rgn21nm), alpha = 0.5) +  # Optional: color points by region if relevant
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(title = "Relationship between undeveloped Land Use and Crime Counts in London",
       x = "Residential Land Use (hectares)", y = "Crime Counts") +
  theme_minimal()

london_data_long <- london_data %>%
  pivot_longer(cols = c(ttl_dv_, ttl_n__), names_to = "LandUseType", values_to = "LandUse") %>%
  mutate(LandUseType = recode(LandUseType, ttl_dv_ = "Developed", ttl_n__ = "Undeveloped"))

# Plotting both regression lines on the same plot
ggplot(london_data_long, aes(x = LandUse, y = count, color = LandUseType)) +
  geom_point(alpha = 0.5) +  # You can still color points by region by adding: aes(group = rgn21nm)
  geom_smooth(method = "lm", se = TRUE, aes(linetype = LandUseType)) +  # Different linetypes for clarity
  labs(title = "Relationship between Land Use and Crime Counts in London",
       x = "Land Use (hectares)", y = "Crime Counts",
       color = "Land Use Type", linetype = "Land Use Type") +
  theme_minimal()

p <- ggplot(london_data, aes(x = ttl_dv_, y = count, text = paste("Region:", rgn21nm))) +
  geom_point(aes(color = rgn21nm), alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(x = "Developed Land Use (hectares)", y = "Crime Counts")
ggplotly(p) %>% layout(title = 'Interactive plot of Developed Land Use vs. Crime Counts')


```

```{r looking specifically at the crime types in london}
# Load necessary libraries
head(london_data_by_crime)

# Assuming london_data_by_crime has columns: LSOA, ttl_dv_, ttl_n__, count, CrimeType
# Pivot longer to get LandUseType and LandUse
london_data_by_crime_long <- london_data_by_crime %>%
  pivot_longer(cols = c(ttl_dv_, ttl_n__), names_to = "LandUseType", values_to = "LandUse") %>%
  mutate(LandUseType = recode(LandUseType, ttl_dv_ = "Developed", ttl_n__ = "Undeveloped"))

# Plotting with facets for each crime type
ggplot(london_data_by_crime_long, aes(x = LandUse, y = count, color = LandUseType)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE, aes(linetype = LandUseType)) +
  labs(title = "Relationship between Land Use and Crime Counts by Crime Type in London",
       x = "Land Use (hectares)", y = "Crime Counts",
       color = "Land Use Type", linetype = "Land Use Type") +
  theme_minimal() +
  facet_wrap(~ crm_typ)  # Facet by crime type

#Focus on only those crime types that have a significant relationship between either the developed or undeveloped land themes

# Remove groups with insufficient data
filtered_data <- london_data_by_crime_long %>%
  group_by(crm_typ, LandUseType) %>%
  filter(!is.na(count) & !is.na(LandUse) & n() > 1) %>%
  ungroup()

# Fit linear models for each CrimeType and LandUseType combination
lm_results <- filtered_data %>%
  group_by(crm_typ, LandUseType) %>%
  nest() %>%
  mutate(model = map(data, ~ lm(count ~ LandUse, data = .x)),
         results = map(model, tidy)) %>%
  unnest(results) %>%
  filter(term == "LandUse") %>%
  select(crm_typ, LandUseType, estimate, p.value)
print(lm_results)


# Filter crime types with significant relationships (p.value < 0.05)
# Filter crime types with significant relationships (p.value < 0.05)
significant_developed <- lm_results %>%
  filter(LandUseType == "Developed" & p.value < 0.05) %>%
  pull(crm_typ)

significant_undeveloped <- lm_results %>%
  filter(LandUseType == "Undeveloped" & p.value < 0.05) %>%
  pull(crm_typ)

significant_crime_types <- lm_results %>%
  filter(p.value < 0.05) %>%
  select(crm_typ) %>%
  distinct()

# Combine significant crime types
significant_crime_types <- union(significant_developed, significant_undeveloped)

# Join significant crime types back with the original data
significant_data <- london_data_by_crime_long %>%
  filter(crm_typ %in% significant_crime_types)

# Plotting with facets for each significant crime type
ggplot(significant_data, aes(x = LandUse, y = count, color = LandUseType)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE, aes(linetype = LandUseType)) +
  labs(title = "Significant Relationships between Land Use and Crime Counts by Crime Type in London",
       x = "Land Use (hectares)", y = "Crime Counts",
       color = "Land Use Type", linetype = "Land Use Type") +
  theme_minimal() +
  facet_wrap(~ crm_typ)  # F


```


```{r hot spot analysis for london }
### spatial weights matrix for london
# Assuming 'full_data' is already read as an sf object
# Compute the spatial weights matrix using contiguity (neighbors sharing boundaries)
neighbors <- poly2nb(london_data)
weights <- nb2listw(neighbors, style="W")

# Moran's I global test
moran_test <- moran.test(london_data$count, weights, zero.policy = TRUE)
print(moran_test)

# Local Moran's I (LISA)
local_moran <- localmoran(london_data$count, weights, zero.policy = TRUE)

# Add LISA results to the data
london_data$LISA <- local_moran[, 1]
london_data$LISA_pvalue <- local_moran[, 5]

# Visualize hotspots using tmap
tm_shape(london_data) +
  tm_polygons("LISA", 
              style = "quantile", 
              palette = "-RdYlBu", 
              title = "Local Moran's I") +
  tm_layout(title = "Hotspot Analysis of Crime Counts by LSOA")
```



```{r spatial lag model for london }
###### Spatial lag model
slm <- lagsarlm(count ~ ttl_dv_ + ttl_n__, data = london_data, listw = weights)
summary(slm)

# Spatial lag model using detailed developed land categories
slm_detailed <- lagsarlm(count ~ grp_cm_ + grp_dfn + grp_nds_ +
                         grp_mn_ + grp_rsd + grp_th_ +
                         grp_tr_, data = london_data, listw = weights)
summary(slm_detailed)

#Group Industry/Commerce (grp_nds_): Each additional hectare is associated with a significant increase in crime (coefficient = 3.691361, p < 2.2e-16), suggesting strong evidence that industrial or commercial land use is associated with higher crime rates.
slm_subgroup <- lagsarlm(count ~ indstry + offices + retail + strg_n_ +            #Group Industry/Commerce (grp_nds_):
                                 resdntl + ins____   +                             #Residential 
                                 hghw___ + trnspr_ + utilits +                     #Transport 
                                 undn___ + undntfd_b + undntfd_s + unknown,        #Other Develeoped Use
                         data = london_data, listw = weights)
summary(slm_subgroup)


summary(slm)
summary(slm_detailed)
summary(slm_subgroup)

```



### Visualise these results 

```{r}
 #1) Map these coordinates 
# Assuming coefficients are stored and spatial data is available
# Calculate predicted values based on one of the models
london_data$predicted_crime = predict(slm, type="response")

# Plotting with tmap
library(tmap)
tmap_mode("plot")
tm_shape(london_data) +
  tm_fill("predicted_crime", title="Predicted Crime Rates", style="quantile", palette="viridis") +
  tm_borders() +
  tm_layout(main.title="Predicted Crime Rates Across London")


### Coeffiecient plot 
coef_table <- summary(slm)$coef
if (is.null(coef_table)) {
  coef_table <- cbind(Estimate = coef(slm), 
                      StdError = sqrt(diag(vcov(slm))),
                      zValue = coef(slm) / sqrt(diag(vcov(slm))))
  
  # Adding p-values
  coef_table <- cbind(coef_table, PValue = 2 * pnorm(-abs(coef_table[, "zValue"])))
}

# Create a dataframe
coef_df <- as.data.frame(coef_table)
coef_df$Variable <- rownames(coef_df)
rownames(coef_df) <- NULL  # Clean up row names

# Display the dataframe
print(coef_df)

library(ggplot2)

# Plotting the coefficients with significance highlighted
ggplot(coef_df, aes(x = Variable, y = Estimate, fill = PValue < 0.05)) +
  geom_col() +
  geom_errorbar(aes(ymin = Estimate - 1.96 * StdError, ymax = Estimate + 1.96 * StdError), width = 0.2) +
  labs(title = "Coefficients of Spatial Lag Model", y = "Coefficient Estimate") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


### Residual Mapping 
# Calculate residuals
london_data$residuals = residuals(slm)

tmap_mode("plot")
tm_shape(london_data) +
  tm_fill("residuals", title="Residuals", style="diverging", midpoint=0, palette="RdBu") +
  tm_layout(main.title="Map of Model Residuals")



###
# Bind coefficient data into a single dataframe
coefficients_all <- rbind(
  transform(as.data.frame(summary_slm$coefficients), Model = "SLM"),
  transform(as.data.frame(summary_slm_detailed$coefficients), Model = "SLM Detailed"),
  transform(as.data.frame(summary_slm_subgroup$coefficients), Model = "SLM Subgroup")
)

coefficients_all <- coefficients_all %>%
  pivot_longer(cols = -c(Model, Estimate, Std.Error, z.value, Pr...z..), names_to = "Variable")

# Plotting
ggplot(coefficients_all, aes(x = Variable, y = Estimate, fill = Model)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_errorbar(aes(ymin = Estimate - 1.96 * `Std.Error`, ymax = Estimate + 1.96 * `Std.Error`), 
                position = position_dodge(width = 0.8), width = 0.25) +
  labs(title = "Comparison of Coefficients Across Models", x = "Variables", y = "Coefficient Estimate") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))





```



```{r spatial error model}
library(sf)
library(sp)
library(spdep)
library(spatialreg)

# Step 1: Load and inspect data
# Ensure full_data is of class 'sf'
class(london_data)

# Step 2: Check and handle missing data
required_vars <- c("count", "indstry", "offices", "retail", "strg_n_", "resdntl", "ins____", 
                   "hghw___", "trnspr_", "utilits", "undn___", "undntfd_b", "undntfd_s", 
                   "unknown")

# Summarize missing data
missing_summary <- sapply(london_data[required_vars], function(x) sum(is.na(x)))
print(missing_summary)

# Handle missing data
# Option 1: Remove rows with missing data
# london_data <- na.omit(london_data)

# Option 2: Impute missing data (example using median imputation)
for (var in required_vars) {
  london_data[[var]][is.na(london_data[[var]])] <- median(london_data[[var]], na.rm = TRUE)
}

# Verify the number of rows
print(nrow(london_data))  # Should match with the original count

# Step 3: Ensure unique identifiers
# If lsoa21nm is not appropriate, create a unique identifier based on row numbers
london_data$unique_id <- as.character(seq_len(nrow(london_data)))

# Step 4: Coerce the sf object into a new sp object
ncovr_s_sp <- as(london_data, "Spatial")

# Ensure row.names match the unique IDs
if (nrow(ncovr_s_sp) == length(london_data$unique_id)) {
  ncovr_s_sp <- spChFIDs(ncovr_s_sp, london_data$unique_id)
} else {
  stop("The lengths of ncovr_s_sp and unique_id do not match.")
}

# Step 5: Create a list of neighbours using the Queen criteria
w_s <- poly2nb(ncovr_s_sp, row.names = row.names(ncovr_s_sp))

# Convert the list of neighbors to a spatial weights matrix
wm_s <- nb2mat(w_s, style = 'B')

# Convert the spatial weights matrix to a listw object
rwm_s <- mat2listw(wm_s, style = 'W')

# Step 6: Fit the spatial error model
fit_3_err <- errorsarlm(count ~ indstry + offices + retail + strg_n_ +            # Group Industry/Commerce (grp_nds_):
                                 resdntl + ins____ +                             # Residential 
                                 hghw___ + trnspr_ + utilits +                   # Transport 
                                 undn___ + undntfd_b + undntfd_s + unknown, 
                                 data = london_data, listw = rwm_s)

# Step 7: Print summary of the model
summary(fit_3_err)


### Plot 
# Residuals plot
plot(fit_3_err$residuals, main = "Residuals Plot", ylab = "Residuals", xlab = "Index")
abline(h = 0, col = "red")

# Add predictions and residuals to the london_data
predicted_values <- predict(fit_3_err, newdata = london_data, listw = rwm_s, type = "response")
residuals_values <- residuals(fit_3_err)

# Add predictions and residuals to the london_data
london_data$predicted <- as.numeric(predicted_values)
london_data$residuals <- residuals_values

# Plot predicted values
ggplot(data = london_data) +
  geom_sf(aes(fill = predicted)) +
  scale_fill_viridis_c() +
  theme_minimal() +
  labs(title = "Predicted Values", fill = "Predicted")

# Plot residuals
ggplot(data = london_data) +
  geom_sf(aes(fill = residuals)) +
  scale_fill_viridis_c() +
  theme_minimal() +
  labs(title = "Residuals", fill = "Residuals")


coefficients <- summary(fit_3_err)$Coef[, 1]
stderr <- summary(fit_3_err)$Coef[, 2]
zvalues <- summary(fit_3_err)$Coef[, 3]
pvalues <- summary(fit_3_err)$Coef[, 4]

coef_df <- data.frame(
  Variable = names(coefficients),
  Estimate = coefficients,
  StdError = stderr,
  zValue = zvalues,
  pValue = pvalues
)

# Remove NA values (e.g., for variables not included due to singularities)
coef_df <- na.omit(coef_df)

# Plot coefficients
ggplot(coef_df, aes(x = Estimate, y = reorder(Variable, Estimate))) +
  geom_point() +
  geom_errorbarh(aes(xmin = Estimate - 1.96 * StdError, xmax = Estimate + 1.96 * StdError), height = 0.2) +
  theme_minimal() +
  labs(title = "Coefficient Plot", x = "Estimate", y = "Variable")

# Load necessary libraries
library(ggplot2)


# Create a dataframe with proper variable names
results <- data.frame(
  Predictor = factor(c("Industry", "Offices", "Retail", "Storage", "Residential", "Institutional", "Transport", "Undefined Built", "Undefined Structure"),
                     levels = c("Industry", "Offices", "Retail", "Storage", "Residential", "Institutional", "Transport", "Undefined Built", "Undefined Structure")),
  Estimate = c(-0.679756, 4.312876, 12.556124, -1.753366, 1.454210, 7.678709, 0.313974, 4.955264, -1.673717)
)

# Create the bar chart
ggplot(results, aes(x = Predictor, y = Estimate, fill = Predictor)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Significant Predictors in the Spatial Error Model", y = "Estimate", x = "Predictor") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


###
# Extract residuals from the model
residuals <- residuals(fit_3_err)

# Assuming 'london_data' has a unique identifier for merging
london_data$residuals <- residuals
ggplot(london_data) +
  geom_sf(aes(fill = residuals)) +
  scale_fill_viridis_c() +
  theme_minimal() +
  labs(title = "Heatmap of Model Residuals", fill = "Residuals")
```




